The association of smoking status with SARS-CoV-2 infection, hospitalisation and mortality from COVID-19: A living rapid evidence review with Bayesian meta-analyses (`r paste(current_version)`)

### 2021-07-27 reports/manuscript.Rmd
v12

Categorisation of smoking status was heterogeneous (see [Table 1](#table-1)). `r paste(data_source[1,2])` `r paste(round(data_source[1,2]/sum(data_source[,2])*100,1), "%", sep = "")` studies collected data on smoking status through routine electronic health records (EHRs), `r paste(data_source[2,2])` `r paste(round(data_source[2,2]/sum(data_source[,2])*100,1), "%", sep = "")` studies used a bespoke case report form for COVID-19 and `r paste(data_source[3,2])` `r paste(round(data_source[3,2]/sum(data_source[,2])*100,1), "%", sep = "")` studies did not state the source for information on smoking status. None of the studies verified smoking status biochemically. Notably, only `r paste(nrow(full_smoking_status))` (`r paste(round(nrow(full_smoking_status)/current_total*100,1))`%) studies reported current, former and never smoking status (see [Supplementary table S2a](#supplementary-t2a)), with a further 26 studies reporting only ever and never smoking status (see [Supplementary table S2b](#supplementary-t2b)). The remaining `r paste(nrow(incomplete_smoking_status))` studies reported current, current/former or current and former smoking status but did not explicitly state whether remaining participants were never smokers or if data were missing on smoking status (see [Supplementary table S2c](#supplementary-t2c)). `r paste(nrow(missing_smok))` studies explicitly reported the proportion with missing data on smoking status, which ranged from `r paste(round(minimum,1), "%", sep = "")` to `r paste(round(maximum,1), "%", sep = "")`.

@@

Fifteen studies recorded the use of alternative nicotine products in current and/or former smokers but did not report COVID-19 outcomes stratified by alternative nicotine use (see [Table 1](#table-1).

@@

Unadjusted smoking prevalence compared with overall estimates for national adult smoking prevalence split by country and study setting is presented in [Figure 2a](#fig-2a) and [2b](#fig-2b). Lower than expected current smoking prevalence was generally observed, especially in studies with hospitalised samples. Former smoking prevalence was more similar to expected prevalence when reported; however, study-based prevalence was typically higher than national estimates. National smoking prevalence estimates used for comparison are presented in [Supplementary table 3](#supplementary-t3).

![*Figure 2.* Weighted mean prevalence of current (left panel) and former (right panel) smoking in included studies, split by country. The circles represent individual studies, with colour corresponding to study setting (i.e. hospital, community and hospital, community) and size corresponding to relative study sample size. For comparison, national current and former smoking prevalence are indicated by the red lines. Countries with three or more eligible studies are shown.](figure/combined_smoking_plots_updated.png){#fig-2a}

@@

`r paste(nrow(table_3_word))` studies examined hospitalisation for COVID-19 disease, stratified by smoking status (see [Table 3](#table-3)). Meta-analyses were performed for one 'good' and 18 ‘fair’ quality studies (see [Figure 5](#fig-5) and [6](#fig-6)). Current (RR = `r paste(round(current_hospitalisation_RR_value, 2))`, 95% CrI = `r paste(round(current_hospitalisation_lci, 2), round(current_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(current_hospitalisation_tau, 2))`) and former (RR = `r paste(round(former_hospitalisation_RR_value, 2))`, CrI = `r paste(round(former_hospitalisation_lci, 2), round(former_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(former_hospitalisation_tau, 2))`) compared with never smokers were at increased risk of hospitalisation with COVID-19. However, data for current smokers were inconclusive and favoured there being no important association. The probability of current and former smokers being at increased risk of hospitalisation (RR ≥1.1) compared with never smokers was `r paste(round(1-m3_a_ecdf(1.1),2)*100,"%", sep = "")` and >99%, respectively. Results were materially unchanged in two sensitivity analyses.

####    [*Table 3* COVID-19 hospitalsation by smoking status]{#table-3}

>>>>

Categorisation of smoking status was heterogeneous (see [Table 1](#table-1)). `r paste(data_source[1,2])` `r paste("(", round(data_source[1,2]/sum(data_source[,2])*100,1), "%",")", sep = "")` studies collected data on smoking status through routine electronic health records (EHRs), `r paste(data_source[2,2])` `r paste("(", round(data_source[2,2]/sum(data_source[,2])*100,1), "%", ")", sep = "")` studies used a bespoke case report form for COVID-19 and `r paste(data_source[3,2])` `r paste( "(", round(data_source[3,2]/sum(data_source[,2])*100,1), "%", ")", sep = "")` studies did not state the source for information on smoking status. None of the studies verified smoking status biochemically. Notably, only `r paste(nrow(full_smoking_status))` (`r paste(round(nrow(full_smoking_status)/current_total*100,1))`%) studies reported current, former and never smoking status (see [Supplementary table S2a](#supplementary-t2a)), with a further 26 studies reporting only ever and never smoking status (see [Supplementary table S2b](#supplementary-t2b)). The remaining `r paste(nrow(incomplete_smoking_status))` studies reported current, current/former or current and former smoking status but did not explicitly state whether remaining participants were never smokers or if data were missing on smoking status (see [Supplementary table S2c](#supplementary-t2c)). `r paste(nrow(missing_smok))` studies explicitly reported the proportion with missing data on smoking status, which ranged from `r paste(round(minimum,1), "%", sep = "")` to `r paste(round(maximum,1), "%", sep = "")`.

@@

Fifteen studies recorded the use of alternative nicotine products in current and/or former smokers but did not report COVID-19 outcomes stratified by alternative nicotine use (see [Table 1](#table-1)).

@@

Unadjusted smoking prevalence compared with overall estimates for national adult smoking prevalence split by country and study setting is presented in [Figure 2](#fig-2). Lower than expected current smoking prevalence was generally observed, especially in studies with hospitalised samples. Former smoking prevalence was more similar to expected prevalence when reported; however, study-based prevalence was typically higher than national estimates. National smoking prevalence estimates used for comparison are presented in [Supplementary table 3](#supplementary-t3).

![*Figure 2.* Weighted mean prevalence of current (left panel) and former (right panel) smoking in included studies, split by country. The circles represent individual studies, with colour corresponding to study setting (i.e. hospital, community and hospital, community) and size corresponding to relative study sample size. For comparison, national current and former smoking prevalence are indicated by the red lines. Countries with three or more eligible studies are shown.](figure/combined_smoking_plots_updated.png){#fig-2}

@@

`r paste(nrow(table_3_word))` studies examined hospitalisation for COVID-19 disease, stratified by smoking status (see [Table 3](#table-3)). Meta-analyses were performed for 19 ‘fair’ quality studies (see [Figure 5](#fig-5) and [6](#fig-6)). Current (RR = `r paste(round(current_hospitalisation_RR_value, 2))`, 95% CrI = `r paste(round(current_hospitalisation_lci, 2), round(current_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(current_hospitalisation_tau, 2))`) and former (RR = `r paste(round(former_hospitalisation_RR_value, 2))`, CrI = `r paste(round(former_hospitalisation_lci, 2), round(former_hospitalisation_uci, 2), sep = "-")`, τ = `r paste(round(former_hospitalisation_tau, 2))`) compared with never smokers were at increased risk of hospitalisation with COVID-19. However, data for current smokers were inconclusive and favoured there being no important association. The probability of current and former smokers being at increased risk of hospitalisation (RR ≥1.1) compared with never smokers was `r paste(round(1-m3_a_ecdf(1.1),2)*100,"%", sep = "")` and >99%, respectively. Results were materially unchanged in two sensitivity analyses.

####    [*Table 3* COVID-19 hospitalisation by smoking status]{#table-3}
